---
title: "Ancona"
author: "Simon Halliday"
date: "2015-November -11"
output: html_document
---

```{r requirements, messages = FALSE}
library(tidyverse)
```

```{r import Ancona}
AnconaBase <- read.csv("/Users/shalliday/Google Drive/github/bfh-textbook/indmarketdemand/ancona.csv")
```


```{r Check Ancona}
head(AnconaBase)
```

```{r data wrangling}
Ancona <- 
  AnconaBase %>%
  rename(date = dTransDataVendita, 
         productid = iProdID, 
         productdescription = vchClasseProdottoDescrizione, 
         speciesname = vchSpecieDescrizione, 
         quantity = nTransQuantita, 
         unitprice = nTransPrezzoUnitario, 
         saleprice = nTransPrezzoVendita, 
         finalprice = nTransPrezzoFinale, 
         producerid = iProduttoreID, 
         qualityid = iQualitaID, 
         qualitydescription = vchQualitaDescrizione, 
         caseid = nNumCasse)
```

```{r}
head(Ancona)
```

```{r averages by date}
AnconaAve <- 
  Ancona %>% 
  group_by(date) %>%
  summarise(meanprice = mean(unitprice), 
            meanquantity = mean(quantity), 
            sumquantity = sum(quantity)/10)
```

```{r}
head(AnconaAve)
```


```{r plot1}
AnconaAve %>% 
  ggplot(aes(x = meanquantity, y = meanprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price") +
  xlab("Daily quantity (kilograms") +
  theme_bw()
```

```{r Plot2}
AnconaAve %>% 
  ggplot(aes(x = sumquantity, y = meanprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw()
```

```{r}
pdf(file = "indmarketdemand/ancona_plot.pdf", width = 5, height = 4)
AnconaAve %>% 
  ggplot(aes(x = sumquantity, y = meanprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw() 
dev.off()
```


```{r}
coef(lm(meanprice ~ sumquantity, data = AnconaAve))
```
I've rounded up the numbers from the above to 9.2 for the intercept and -0.0006 for the slope. 

```{r}
pdf(file = "indmarketdemand/ancona_plot2.pdf", width = 5, height = 4)
AnconaAve %>% 
  ggplot(aes(x = sumquantity, y = meanprice)) +
  geom_point() +
  geom_abline(intercept = 9.2, slope = -0.0006, col = "blue", cex = 0.9) +
  ylab(expression(paste("Daily average price per kilogram (Euros), ", p[x]))) +
  xlab("Daily quantity (kilograms) of fish, x") +
  scale_x_continuous(limits = c(0,1600), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0,12), expand = c(0, 0)) +
  annotate("text", x = 1000, y = 5.5, label = "Marginal Rate of") +
  annotate("text", x = 1000, y = 5, label = "Substitution, or") +
  annotate("text", x = 1000, y = 4.5, label = "Demand for x") +
  annotate("text", x = 53, y = 9.4, label = "r[max]", parse = TRUE) +
  annotate("text", x = 1530, y = 3, label = "x[max]", parse = TRUE) +
  geom_segment(aes(x = 1530, y = 2.7, xend = 1530, yend = 0.2), 
               size = 0.7, arrow = arrow(type = "closed", 
                                         length = unit(0.25, "cm"), 
                                         angle = 25)) +
  theme_bw() 
dev.off()
```


```{r}
m1 <- lm(sumquantity ~ meanprice, data = AnconaAve)
summary(m1)
m2 <- lm(meanprice ~ sumquantity, data = AnconaAve)
summary(m2)
```


```{r natural logs}
AnconaAve2 <- 
  AnconaAve %>%
  mutate(lnprice = log(meanprice), lnquantity = log(sumquantity))
AnconaAve2 %>% 
  ggplot(aes(x = sumquantity, y = lnprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw()
AnconaAve2 %>% 
  ggplot(aes(x = lnquantity, y = meanprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw()
```

Now using 'codiceacquirente' - I think that might be buyerid...


```{r Buyer 78 v2}
Ancona78v3 <- 
  Ancona %>%
  filter(CodiceAcquirente == 78)
head(Ancona78v3)
```

```{r}
Ancona78v3Aves <- 
  Ancona78v3 %>%
  group_by(date) %>%
  summarise(dailyquantity = sum(quantity), dailyaveprice = mean(unitprice), meanquantity = mean(quantity))
```

```{r}
Ancona78v3Aves %>%
    ggplot(aes(x = dailyquantity, y = dailyaveprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw()
```

OK... Wait. Crucially, they say in the text (book, page 15) that the graph is only for those values of kilograms less than 200. 

```{r}
Ancona78v3Aves %>%
  filter(dailyquantity < 200) %>%
    ggplot(aes(x = dailyquantity, y = dailyaveprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw()
```

OK. Now export this as a pdf.



```{r Buyer78 Graph}
pdf(file = "bfh-textbook/indmarketdemand/ancona_buyer78.pdf", width = 5, height = 4)
Ancona78v3Aves %>%
  #filter(dailyquantity < 200) %>%
    ggplot(aes(x = dailyquantity, y = dailyaveprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw() +
  theme(panel.grid=element_blank(), panel.border=element_blank())
dev.off()
```





```{r}
Ancona78v3Aves %>%
  filter(dailyquantity < 400) %>%
    ggplot(aes(x = dailyquantity, y = dailyaveprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw()
```







```{r Buyer 78}
Ancona78 <- 
  Ancona %>%
  filter(producerid == 78)
head(Ancona78)
```

```{r}
Ancona78 %>%
    ggplot(aes(x = quantity, y = unitprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw() + 
  theme(panel.grid=element_blank(), panel.border=element_blank())
```
This doesn't match the graphic they have in the book for buyer 78. I may have filtered on the wrong variable. 

OK. I checked the Gallegati et al paper and it looks like I needed it to be averages (we didn't have this mentioned in our draft). 

```{r Buyer 78 Averages}
Ancona78Aves <- 
  Ancona78 %>%
  group_by(date) %>%
  summarise(dailyquantity = sum(quantity), dailyaveprice = mean(unitprice), meanquantity = mean(quantity))
```

```{r}
Ancona78Aves %>%
    ggplot(aes(x = dailyquantity, y = dailyaveprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw()
```


```{r Buyer 78 v2}
Ancona78v2 <- 
  Ancona %>%
  filter(productid == 78)
head(Ancona78v2)
```

```{r}
Ancona78v2Aves <- 
  Ancona78v2 %>%
  group_by(date) %>%
  summarise(dailyquantity = sum(quantity), dailyaveprice = mean(unitprice), meanquantity = mean(quantity))
```

```{r}
Ancona78v2Aves %>%
    ggplot(aes(x = dailyquantity, y = dailyaveprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw()
```



And now attempt a ggplot of this data: 
```{r}
Ancona78v2 %>%
    ggplot(aes(x = quantity, y = unitprice)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE) + 
  ylab("Daily average price per kilogram (Euros)") +
  xlab("Daily quantity (kilograms)") +
  theme_bw()
```




