---
title: "FRED 30Y Fixed Rate Mortgage x Housing Starts"
author: "Scott Cohn"
date: "June 22"
output: 
    html_document:
        theme: flatly
        highlight: haddock
        toc: yes
        toc_depth: 4
        toc_float: no 
        keep_md: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, cache = TRUE)

#' GDPDEF: https://fred.stlouisfed.org/series/GDPDEF (annual)
#' M30: https://fred.stlouisfed.org/series/MORTGAGE30US (quarterly)
#' HOUST: https://fred.stlouisfed.org/series/HOUST (quarterly)

library("tidyverse")
library("zoo")

library("sjPlot")
library("sjmisc")
library("sjlabelled")

library("ggpubr")

```

# Clean + Tidy + Mutate
```{r clean}

# load gdpdeflator
df_gdpdef <- read.csv("fred_data/GDPDEF.csv")

# convert date + inflation
df_gdpdef$GDPDEF <- as.numeric(df_gdpdef$GDPDEF)

df_gdpdef <- df_gdpdef %>%
  mutate(DATE = as.Date(df_gdpdef$DATE, format = "%Y-%m-%d"),
         inflation = ((GDPDEF - lag(GDPDEF, default = 0))/lag(GDPDEF, default = 0))*100)

# change inf
df_gdpdef[1, 3] <- 0

# load m30 and houst
df <- read_csv("fred_data/m30_x_houst.csv")

# merge on date
df <- right_join(df_gdpdef, df, by = "DATE")

# carry last NA on gdpdef
df <- df %>%
  mutate(
    GDPDEF = na.locf(df$GDPDEF),
    inflation = na.locf(df$inflation)
  )

# rename
df <- df %>%
  rename(c("houst" = "HOUST", "m30" = "MORTGAGE30US"))

# as numeric
df <- df %>% 
  mutate(houst = as.numeric(df$houst),
         m30 = as.numeric(df$m30))

# real m30 (rm30)
df <- df %>% mutate(rm30 = m30 - inflation)

```

```{r lag_calcs, echo=FALSE}
# moving avg
df <- df %>% mutate(rm30_ma_k3 = rollmean(rm30, k = 3, fill = NA),
                    rm30_ma_k6 = rollmean(rm30, k = 6, fill = NA),
                    rm30_ma_k9 = rollmean(rm30, k = 9, fill = NA))

# lagged
df <- df %>%
  mutate(rm30_lag2q_ma_k3 = lag(df$rm30_ma_k3, 2, na.pad = TRUE), # 6month
         rm30_lag3q_ma_k3 = lag(df$rm30_ma_k3, 3, na.pad = TRUE), # 9month
         rm30_lag4q_ma_k3 = lag(df$rm30_ma_k3, 4, na.pad = TRUE), # 12month
         rm30_lag2q_ma_k6 = lag(df$rm30_ma_k6, 2, na.pad = TRUE), # 6month
         rm30_lag3q_ma_k6 = lag(df$rm30_ma_k6, 3, na.pad = TRUE), # 9month
         rm30_lag4q_ma_k6 = lag(df$rm30_ma_k6, 4, na.pad = TRUE), # 12month
         rm30_lag2q_ma_k9 = lag(df$rm30_ma_k9, 2, na.pad = TRUE), # 6month
         rm30_lag3q_ma_k9 = lag(df$rm30_ma_k9, 3, na.pad = TRUE), # 9month
         rm30_lag4q_ma_k9 = lag(df$rm30_ma_k9, 4, na.pad = TRUE)) # 12month

```

# Plots

### houst x date

```{r houst_date, echo=FALSE}

df %>% 
  gather(metric, value, rm30:rm30_lag4q_ma_k9) %>%
  filter(DATE > "1970-01-01" & DATE < "2006-01-01") %>% 
  ggplot(aes(DATE, houst)) +
  geom_line() +
  theme_bw() +
  labs(x = "Date", y = "Housing Starts: Total: New Privately Owned Housing Units Started", caption = "Source: FRED")

#ggsave("FRED_houst.pdf", width = 9, height = 6)

```

### houst x rm30 (1yrlag)

```{r houst_rm30_1yrlag, echo=FALSE}

df %>% 
  select(DATE, houst, rm30) %>% 
  # lagged diffs (1 year // 4Qs)
  mutate(houst_lag_diff = houst - lag(df$houst, 4, na.pad = TRUE),
         rm30_lag_diff = rm30 - lag(df$rm30, 4, na.pad = TRUE)) %>% 
  #filter(DATE > "1977-01-01" & DATE < "1985-01-01") %>% 
  filter(DATE > "1977-01-01" & DATE < "2006-01-01") %>% 
  ggplot(aes(x = rm30_lag_diff, y = houst_lag_diff)) + 
  geom_point(color = "#377EB8") +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Real 30-year fixed rate mortgage average  differences (1 Year)", 
       y = "Housing start differences (1 Year)",
       caption = "1977-2006") + 
  theme_bw() + 
  theme(panel.grid.minor = element_blank(),
        axis.title = element_text(size = 17),
        axis.text = element_text(size = 12))

#ggsave("FRED_houst_rm30_first_differences_77_06.pdf", width = 9, height = 6)

```

### houst x rm30

```{r houst_rm30, echo=FALSE}

df %>% 
  select(DATE, houst, rm30) %>% 
  filter(DATE > "1970-01-01" & DATE < "2006-01-01") %>% 
  ggplot(aes(x = rm30, y = houst)) + 
  geom_point(color = "#377EB8") +
  scale_color_brewer(palette = "Set1") +
  theme_bw() + 
  labs(x = "rm30", y = "houst")

#ggsave("FRED_houst_rm30.pdf", width = 9, height = 6)

```

### m30 x date

```{r m30_date, echo=FALSE}

df %>% 
  # wide to long
  gather(metric, value, inflation:rm30_ma_k9) %>%
  #filter(metric %in% c("inflation", "m30", "rm30"),
  filter(metric != "houst",
    DATE > "1970-01-01" & DATE < "2006-01-01") %>% 
  # plot
  ggplot(aes(DATE, value, color = metric)) +
  geom_line() +
  scale_color_brewer(palette = "Set1") + 
  theme_bw() + 
  labs(x = "Date", y = "Percentage", caption = expression(paste("Source: FRED")))

#ggsave("FRED_m30.pdf", width = 9, height = 6)

```

### Dual axis 

Ignore this for now.

```{r dual_axis, echo=FALSE}

coeff <- 210
df %>% 
  gather(metric, value, rm30_lag2q_ma_k3:rm30_lag4q_ma_k9) %>%
  filter(DATE > "1970-01-01" & DATE < "2006-01-01") %>% 
  ggplot(aes(x = DATE)) +
  #ggplot(aes(DATE, houst)) +
  geom_line(aes(y=value, color = metric)) +
  geom_line(aes(y=houst/coeff), color = "black") +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Percentage",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*coeff, name="Housing Starts")
  ) + 
  #scale_fill_manual(values = colorRampPalette(13)) +
  #scale_color_brewer(palette = "Set3") +
  #labs(x = "Date", y = "Housing Starts: Total: New Privately Owned Housing Units Started", caption = "Source: FRED")
  scale_color_brewer(palette = "Set1") + 
  theme_bw() + 
  labs(x = "Date", caption = expression(paste("Source: FRED")))

#ggsave("FRED_dualaxis.pdf", width = 9, height = 6)

```


# Lagged Correlations (1977 -- 2006) 

```{r df_lag_diffs}

df_lag_diffs <- df %>% 
  select(DATE, houst, rm30) %>% 
  # lagged diffs (1 year // 4Qs)
  mutate(houst_lag_diff_1q = houst - lag(df$houst, 1, na.pad = TRUE),    # 1Q
         houst_lag_diff_2q = houst - lag(df$houst, 2, na.pad = TRUE),    # 2Q
         houst_lag_diff_3q = houst - lag(df$houst, 3, na.pad = TRUE),
         houst_lag_diff_4q = houst - lag(df$houst, 4, na.pad = TRUE),
         houst_lag_diff_5q = houst - lag(df$houst, 5, na.pad = TRUE),
         houst_lag_diff_6q = houst - lag(df$houst, 6, na.pad = TRUE),
         houst_lag_diff_7q = houst - lag(df$houst, 7, na.pad = TRUE),
         houst_lag_diff_8q = houst - lag(df$houst, 8, na.pad = TRUE),
         houst_lag_diff_9q = houst - lag(df$houst, 9, na.pad = TRUE),
         
         rm30_lag_diff_1q = rm30 - lag(df$rm30, 1, na.pad = TRUE),    # 1Q
         rm30_lag_diff_2q = rm30 - lag(df$rm30, 2, na.pad = TRUE),    # 2Q
         rm30_lag_diff_3q = rm30 - lag(df$rm30, 3, na.pad = TRUE),
         rm30_lag_diff_4q = rm30 - lag(df$rm30, 4, na.pad = TRUE),
         rm30_lag_diff_5q = rm30 - lag(df$rm30, 5, na.pad = TRUE),
         rm30_lag_diff_6q = rm30 - lag(df$rm30, 6, na.pad = TRUE),
         rm30_lag_diff_7q = rm30 - lag(df$rm30, 7, na.pad = TRUE),
         rm30_lag_diff_8q = rm30 - lag(df$rm30, 8, na.pad = TRUE),
         rm30_lag_diff_9q = rm30 - lag(df$rm30, 9, na.pad = TRUE)
         ) %>% 
  #filter(DATE > "1977-01-01" & DATE < "1985-01-01") %>% 
  filter(DATE > "1977-01-01" & DATE < "2006-01-01") 
  
```

## First Differences: Simple OLS of diff starts on diff mortgages (by quarter)

1 year lag for `houst` regressed on 1 year lag for `rm30`. 

```{r ols_simple_1y_diff, fig.align='center'}

ols_simple_1y <- lm(houst_lag_diff_4q ~ rm30_lag_diff_4q, data = df_lag_diffs)
#summary(ols_simple_1y)
tab_model(ols_simple_1y)
```

Check for stationarity of var:

```{r acf_houst, fig.align='center'}

acf(as.ts(df$houst), na.action = na.pass, plot = TRUE)
```

```{r acf_rm30, fig.align='center'}
acf(as.ts(df$rm30), na.action = na.pass, plot = TRUE)
```

Check for autocorrelation:

```{r ols_1y_autocorr}
res <- ols_simple_1y$res 
n <- length(res) 
ols_1y_autocorr <- lm(res[-n] ~ res[-1]) 
# summary(ols_1y_autocorr)
tab_model(ols_1y_autocorr)

```

Since the coefficient for `res[-1]` is significant, we have evidence for autocorrelation.

## diff starts on diff mortages {.tabset .tabset-fade .tabset-pills}
t-1, t-2, t-3, t-4, ... t-8  (where the numbers refer to quarters (look at the pattern of the coefficients on the lagged mortages)

### 1Q

1 quarter lag for `houst` regressed on 1 quarter lag for `rm30`. 

```{r ols_simple_1q_diff, fig.align='center'}

ols_simple_1q_diff_diff <- lm(houst_lag_diff_1q ~ rm30_lag_diff_1q, data = df_lag_diffs)
#summary(ols_simple_1q_diff_diff)
tab_model(ols_simple_1q_diff_diff)
```

### 2Q

2 quarter lag for `houst` regressed on 2 quarter lag for `rm30`. 

```{r ols_simple_2q_diff, fig.align='center'}

ols_simple_2q_diff_diff <- lm(houst_lag_diff_2q ~ rm30_lag_diff_2q, data = df_lag_diffs)
#summary(ols_simple_2q_diff_diff)
tab_model(ols_simple_2q_diff_diff)
```

### 3Q

3 quarter lag for `houst` regressed on 3 quarter lag for `rm30`. 

```{r ols_simple_3q_diff, fig.align='center'}

ols_simple_3q_diff_diff <- lm(houst_lag_diff_3q ~ rm30_lag_diff_3q, data = df_lag_diffs)
#summary(ols_simple_3q_diff_diff)
tab_model(ols_simple_3q_diff_diff)
```

### 4Q

4 quarter lag for `houst` regressed on 4 quarter lag for `rm30`. 

```{r ols_simple_4q_diff, fig.align='center'}

ols_simple_4q_diff <- lm(houst_lag_diff_4q ~ rm30_lag_diff_4q, data = df_lag_diffs)
#summary(ols_simple_4q_diff)
tab_model(ols_simple_4q_diff)
```

### 5Q

5 quarter lag for `houst` regressed on 5 quarter lag for `rm30`. 

```{r ols_simple_5q_diff, fig.align='center'}

ols_simple_5q_diff <- lm(houst_lag_diff_5q ~ rm30_lag_diff_5q, data = df_lag_diffs)
#summary(ols_simple_5q_diff)
tab_model(ols_simple_5q_diff)
```

### 6Q

6 quarter lag for `houst` regressed on 6 quarter lag for `rm30`. 

```{r ols_simple_6q_diff, fig.align='center'}

ols_simple_6q_diff <- lm(houst_lag_diff_6q ~ rm30_lag_diff_6q, data = df_lag_diffs)
#summary(ols_simple_6q_diff)
tab_model(ols_simple_6q_diff)
```

### 7Q

7 quarter lag for `houst` regressed on 7 quarter lag for `rm30`. 

```{r ols_simple_7q_diff, fig.align='center'}

ols_simple_7q_diff <- lm(houst_lag_diff_7q ~ rm30_lag_diff_7q, data = df_lag_diffs)
#summary(ols_simple_7q_diff)
tab_model(ols_simple_7q_diff)
```

### 8Q

8 quarter lag for `houst` regressed on 8 quarter lag for `rm30`. 

```{r ols_simple_8q_diff, fig.align='center'}

ols_simple_8q_diff <- lm(houst_lag_diff_8q ~ rm30_lag_diff_8q, data = df_lag_diffs)
#summary(ols_simple_8q_diff)
tab_model(ols_simple_8q_diff)
```

### 9Q

9 quarter lag for `houst` regressed on 9 quarter lag for `rm30`. 

```{r ols_simple_9q_diff, fig.align='center'}

ols_simple_9q_diff <- lm(houst_lag_diff_9q ~ rm30_lag_diff_9q, data = df_lag_diffs)
#summary(ols_simple_9q_diff)
tab_model(ols_simple_9q_diff)
```


## starts (not diffs) on various lagged mortages (not diffs) with lags as above. {.tabset .tabset-fade .tabset-pills}

```{r df_lag_nondiff}
df_lag <- df %>% 
  select(DATE, houst, rm30) %>% 
  # lagged diffs (1 year // 4Qs)
  mutate(rm30_lag_1q = lag(df$rm30, 1, na.pad = TRUE),    # 1Q
         rm30_lag_2q = lag(df$rm30, 2, na.pad = TRUE),    # 2Q
         rm30_lag_3q = lag(df$rm30, 3, na.pad = TRUE),
         rm30_lag_4q = lag(df$rm30, 4, na.pad = TRUE),
         rm30_lag_5q = lag(df$rm30, 5, na.pad = TRUE),
         rm30_lag_6q = lag(df$rm30, 6, na.pad = TRUE),
         rm30_lag_7q = lag(df$rm30, 7, na.pad = TRUE),
         rm30_lag_8q = lag(df$rm30, 8, na.pad = TRUE),
         rm30_lag_9q = lag(df$rm30, 9, na.pad = TRUE)
         ) %>% 
  #filter(DATE > "1977-01-01" & DATE < "1985-01-01") %>% 
  filter(DATE > "1977-01-01" & DATE < "2006-01-01") 
```

### 1Q

`houst` regressed on 1 quarter lag for `rm30` .  

```{r ols_simple_1q, fig.align='center'}

ols_simple_1q <- lm(houst ~ rm30_lag_1q, data = df_lag)
#summary(ols_simple_1q)
tab_model(ols_simple_1q)
```

### 2Q

`houst` regressed on 2 quarter lag for `rm30` . 

```{r ols_simple_2q, fig.align='center'}

ols_simple_2q <- lm(houst ~ rm30_lag_2q, data = df_lag)
#summary(ols_simple_2q)
tab_model(ols_simple_2q)
```

### 3Q

`houst` regressed on 3 quarter lag for `rm30` . 

```{r ols_simple_3q, fig.align='center'}

ols_simple_3q <- lm(houst ~ rm30_lag_3q, data = df_lag)
#summary(ols_simple_3q)
tab_model(ols_simple_3q)
```

### 4Q

`houst` regressed on 4 quarter lag for `rm30` . 

```{r ols_simple_4q, fig.align='center'}

ols_simple_4q <- lm(houst ~ rm30_lag_4q, data = df_lag)
#summary(ols_simple_4q)
tab_model(ols_simple_4q)
```

### 5Q

`houst` regressed on 5 quarter lag for `rm30`.

```{r ols_simple_5q, fig.align='center'}

ols_simple_5q <- lm(houst ~ rm30_lag_5q, data = df_lag)
#summary(ols_simple_5q)
tab_model(ols_simple_5q)
```

### 6Q

`houst` regressed on 6 quarter lag for `rm30` . 

```{r ols_simple_6q, fig.align='center'}

ols_simple_6q <- lm(houst ~ rm30_lag_6q, data = df_lag)
#summary(ols_simple_6q)
tab_model(ols_simple_6q)
```

### 7Q

`houst` regressed on 7 quarter lag for `rm30` . 

```{r ols_simple_7q, fig.align='center'}

ols_simple_7q <- lm(houst ~ rm30_lag_7q, data = df_lag)
#summary(ols_simple_7q)
tab_model(ols_simple_7q)
```

### 8Q

`houst` regressed on 8 quarter lag for `rm30`. 

```{r ols_simple_8q, fig.align='center'}

ols_simple_8q <- lm(houst ~ rm30_lag_8q, data = df_lag)
#summary(ols_simple_8q)
tab_model(ols_simple_8q)
```

### 9Q

9 quarter lag for `rm30` regressed on 9 quarter lag for `houst`. 

```{r ols_simple_9q, fig.align='center'}

ols_simple_9q <- lm(houst ~ rm30_lag_9q, data = df_lag)
#summary(ols_simple_9q)
tab_model(ols_simple_9q)
```

# Book Figs

## houst x lagged rm30 (4Q)

```{r bookfig_a, fig.align='center'}

coeff <- 210
df_lag %>% 
  select(DATE, houst, rm30_lag_4q) %>% 
  # gather(metric, value, 2:3) %>%
  filter(DATE > "1970-01-01" & DATE < "2006-01-01") %>% 
  ggplot(aes(x = DATE)) +
  geom_line(aes(y=houst, color = "houst")) +
  geom_line(aes(y=rm30_lag_4q*coeff, color = "rm30_lag_4q")) +
  scale_y_continuous(
    
    # Features of the first axis
    name = "Housing starts: total",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./coeff, name="Real 30-year fixed rate mortgage averages, 1-year lag (%)")
  ) + 
  labs(x = "Date") +
  scale_color_manual(name = "", labels = c("Housing starts", "Real mortgage rates (1-year lag)"), 
                     values = c("#377EB8", "#4DAF4A")) +
  theme_bw() + 
  theme(panel.grid.minor = element_blank(),
        axis.title = element_text(size = 17),
        axis.text = element_text(size = 14),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.925),
        legend.text = element_text(size = 12),
        axis.text.y.right = element_text(margin = unit(c(5, 5, 5, 1), "mm")))

#ggsave("FRED_dualaxis.pdf", width = 9, height = 6)

# df_lag %>% 
#   select(DATE, houst, rm30, rm30_lag_4q) %>% 
#   ggplot(aes(x = rm30_lag_4q, y = houst)) + 
#   geom_point(color = "#377EB8") +
#   labs(x = "Real 30-year fixed rate mortgage averages, 1-year lag", 
#        y = "Housing starts: total") +
#   theme_bw() + 
#   theme(panel.grid.minor = element_blank(),
#         axis.title = element_text(size = 15),
#         axis.text = element_text(size = 13))

ggsave("rm30_lag_4q_x_houst.pdf", width = 9, height = 7, units = "in")
```

## scatter of rm30 diffs x houst diffs

```{r bookfig_b, fig.align='center'}

# cor(df_lag_diffs$houst_lag_diff_4q, df_lag_diffs$rm30_lag_diff_4q) 
# = -0.3974753

summary(lm(houst_lag_diff_4q ~ rm30_lag_diff_4q, data = df_lag_diffs))

df_lag_diffs %>% 
  select(DATE, rm30_lag_diff_4q, houst_lag_diff_4q) %>% 
  ggplot(aes(x = rm30_lag_diff_4q, y = houst_lag_diff_4q)) + 
  geom_smooth(method='lm', se = FALSE, color = "#E41A1C", size = 0.5) +
  # geom_abline(intercept = 14.19, slope = -77.45, color = "#E41A1C", size = 0.5) +
  geom_point(color = "#377EB8") +
  # annotate("segment", x = 4.5, xend = 4.5, y = -500, yend = -350, size=0.5, alpha=1, arrow = arrow(length = unit(0.15, "cm"), type = "closed")) +
  # annotate("text", x = 4.5, y = -540, label = expression(r == 0.397), size = 5) +
  # annotate("text", x = 4.5, y = -600, 
  #          label = expression(paste("Slope ", phantom() == - 77.45)), size = 5) +
  #   annotate("text", x = 4.5, y = -660, 
  #          label = expression(paste("CI", phantom() == phantom(), "(-110.8, -44.13)")), size = 5) +
  # stat_cor(method = "pearson", r.digits = 3, p.digits = 3, label.x = 3.5, label.y = -540) +
  labs(x = "Change in inflation adjusted mortgage interest rate (lagged 1-year)",
      #x = "Real 30-year fixed rate mortgage averages, 1-year lag (differences)", 
      #y = "Housing starts: total (differences)") +
      y = "Change in private housing starts (in thousands)") +
  theme_bw() + 
  theme(panel.grid.minor = element_blank(),
        axis.title = element_text(size = 19),
        axis.text = element_text(size = 17))

ggsave("rm30_lagdiff4q_x_houst_lagdiff4q.pdf", width = 9, height = 7, units = "in")
```

```{r faux_elas}
# Calculate faux elasticity

df_faux_elas <- df_lag %>% 
  select(DATE, houst, rm30_lag_4q) %>%  
  mutate(houst_lag_4q = lag(df_lag$houst, 4, na.pad = TRUE)) %>% 
  filter(DATE > "1978-01-01" & DATE < "2006-01-01") 

w <- mean(df_faux_elas$rm30_lag_4q) / mean(df_faux_elas$houst_lag_4q) 

fe <- w*(-77.45)

fe
```

```{r r_ci}
# calculate CI for pearson r
n = 111 
r = 0.397

stderr <- 1 / sqrt(n-3)

r_lb <- formatC(r - stderr)
r_ub <- formatC(r + stderr)

print(paste0("(", r_lb, ", ", r_ub, " )"))
```
